Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-03-13 Grassmannian_cells.tex
added argument to proof of 4.18, minor changes

The results of \cite{rupel} giving a combinatorial construction of counting polynomials for quiver Grassmannians of preprojective/preinjective representations of generalized Kronecker quivers $K(n)$ suggest that the dimensions of cells can be computed using this combinatorics (or the equivalent combinatorics of compatible pairs).
This is made precise in Conjecture~\ref{conj:cell dimensions}.
\sayT{maybe we can erase some parts which show up twice in the introduction?}
\subsection*{Acknowledgements}We would like to thank Giovanni Cerulli Irelli, Hans Franzen, Oliver Lorscheid and Markus Reineke for very fruitful discussions related to this project.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiver Covering Theory}
\label{sec:covering}
\noindent
We refer to \cite{gab} for an introduction to covering theory.
Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:s(\alpha)\to t(\alpha)$.
A $\CC$-representation $X$ of $Q$ consists of a collection of $\CC$-vector spaces $X_i$ for $i\in Q_0$ and a collection of $\CC$-linear maps $X_\alpha:X_{s(\alpha)}\to X_{t(\alpha)}$ for $\alpha\in Q_1$.
Given $\CC$-representations $X$ and $Y$ of $Q$, a morphism $f:X\to Y$ is a collection of $\CC$-linear maps $f_i:X_i\to Y_i$ for $i\in Q_0$ satisfying $f_{t(\alpha)}\circ X_\alpha=Y_\alpha\circ f_{s(\alpha)}$ for each $\alpha\in Q_1$.
We write $\rep Q$ for the hereditary abelian category of finite-dimensional $\C$-representations of $Q$ and we assume in the following that all representations are finite-dimensional.

@@

\begin{corollary}
  \label{fpoly}
  Let $X$ be a representation which can be lifted to the universal covering quiver.
  \begin{enumerate}
    \item If $\Gr_\bfe(X)$ is smooth and $\Gr^{\tilde Q}_\tbfe(\tilde X)$ has a cell decomposition into affine spaces, then $\Gr_\bfe(X)$ has a cell decomposition into affine spaces.
    \item We have $F_X=SF_{\tilde X}$ where $SF_{\tilde X}$ is obtained from $F_{\tilde X}$ by applying $S:\ZZ[y_{(i,w)}\mid i\in Q_0,w\in W_Q]\to\ZZ[y_i\mid i\in Q_0]$ given by $S(y_{(i,w)})=y_i$ for all $i\in Q_0$ and $w\in W_Q$.
  \end{enumerate} 
\end{corollary}
An important special case for this is the case of exceptional representations.
In this case the quiver Grassmannians $\Gr_\bfe(X)$ are smooth by \cite[Corollary 4]{cr}.
Moreover, every exceptional representation is a tree module by \cite{rin1} which means that it can be lifted to the universal covering.

@@

\begin{theorem}
  \label{celldec}
  The following hold:
  \begin{enumerate}
    \item Every quiver Grassmannian of any indecomposable preprojective or preinjective representation of $K(n)$ and $\widetilde{K(n)}$ has a cell decomposition into affine spaces.
    \item Let $X\in\rep K(n)$ be an indecomposable representation with dimension vector $(d,e)$ or $(e,d)$, where $(d,e)=\udim P_{m+1}-r\udim P_m$ for $m\geq 1$ and $\leq 0\leq r\leq n-1$.
      Then every quiver Grassmannian $\Gr_\bfe(X)$ has a cell decomposition into affine spaces.
  \end{enumerate}
\end{theorem}
\begin{proof}
  The first part follows by combining the results of Section~\ref{torusaction} and Theorems~\ref{thm:bb} and~\ref{cellscover}, taking into account that every preprojective representation of $\widetilde{ K(n)}$ is a lift of a preprojective representation (c.f.\ Lemma~\ref{le:lifts of transjectives}).
  The dual version for preinjective representations follows immediately since $\Gr_{\bfe}(P_m)\cong \Gr_{\udim P_m-\bfe}(I_m)$ and $\Gr_{\tbfe}(\tilde P_m)\cong\Gr_{\udim \tilde P_m-\tbfe}(\tilde I_m)$.

  The second part follows in the same way, additionally, taking Theorem~\ref{thm:truncpp} into account.
\end{proof}

@@

\begin{remark}
  Below we will usually apply Lemma~\ref{lem:2equivalence} after performing each of the equivalences from Definition~\ref{def:2equivalence}.
  That is, given a 2-arrow $V:\Gamma(1)\implies\Gamma(2)$ for which $\Gamma(1)$ has a unique source $p$ and $\Gamma(2)$ has a unique sink $q$, we get an equivalent 2-quiver by replacing this 2-arrow with a usual arrow $a_V:p\to q$.
\end{remark}

>>>>

The results of \cite{rupel} giving a combinatorial construction of counting polynomials for quiver Grassmannians of preprojective/preinjective representations of generalized Kronecker quivers $K(n)$ suggest that the dimensions of cells can be computed using this combinatorics (or the equivalent combinatorics of compatible pairs).
This is made precise in Conjecture~\ref{conj:cell dimensions}.
%\sayT{maybe we can erase some parts which show up twice in the introduction?}
\subsubsection*{Acknowledgements}We would like to thank Giovanni Cerulli Irelli, Hans Franzen, Oliver Lorscheid and Markus Reineke for very fruitful discussions related to this project.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiver Covering Theory}
\label{sec:covering}
\noindent
We refer to \cite{gab} for an introduction to covering theory.
Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:s(\alpha)\to t(\alpha)$.
A $\CC$-representation $X$ of $Q$ consists of a collection of $\CC$-vector spaces $X_i$ for $i\in Q_0$ and a collection of $\CC$-linear maps $X_\alpha:X_{s(\alpha)}\to X_{t(\alpha)}$ for $\alpha\in Q_1$.
Given $\CC$-representations $X$ and $Y$ of $Q$, a morphism $f:X\to Y$ is a collection of $\CC$-linear maps $f_i:X_i\to Y_i$ for $i\in Q_0$ satisfying $f_{t(\alpha)}\circ X_\alpha=Y_\alpha\circ f_{s(\alpha)}$ for each $\alpha\in Q_1$.
We write $\rep Q$ for the hereditary abelian category of finite-dimensional $\C$-representations of $Q$ and we assume in the following that all representations are finite-dimensional.

@@

\begin{corollary}
  \label{fpoly}
  Let $X$ be a representation which can be lifted to the universal covering quiver.
  \begin{enumerate}
    \item If $\Gr_\bfe(X)$ is smooth and $\Gr^{\hat Q}_\tbfe(\hat X)$ has a cell decomposition into affine spaces, then $\Gr_\bfe(X)$ has a cell decomposition into affine spaces.
    \item We have $F_X=SF_{\tilde X}$ where $SF_{\tilde X}$ is obtained from $F_{\tilde X}$ by applying $S:\ZZ[y_{(i,w)}\mid i\in Q_0,w\in W_Q]\to\ZZ[y_i\mid i\in Q_0]$ given by $S(y_{(i,w)})=y_i$ for all $i\in Q_0$ and $w\in W_Q$.
  \end{enumerate} 
\end{corollary}
An important special case for this is the case of exceptional representations.
In this case the quiver Grassmannians $\Gr_\bfe(X)$ are smooth by \cite[Corollary 4]{cr}.
Moreover, every exceptional representation is a tree module by \cite{rin1} which means that it can be lifted to the universal covering.

@@

\begin{theorem}
  \label{celldec}
  The following hold:
  \begin{enumerate}
    \item Every quiver Grassmannian of any indecomposable preprojective or preinjective representation of $K(n)$ and $\widetilde{K(n)}$ has a cell decomposition into affine spaces.
    \item Let $X\in\rep K(n)$ be an indecomposable representation with dimension vector $(d,e)$ or $(e,d)$, where $(d,e)=\udim P_{m+1}-r\udim P_m$ for $m\geq 1$ and $\leq 0\leq r\leq n-1$.
      Then every quiver Grassmannian $\Gr_\bfe(X)$ has a cell decomposition into affine spaces.
  \end{enumerate}
\end{theorem}
\begin{proof}Both claims follow by considering iterated torus actions taking into account that all quiver Grassmannians under consideration are smooth. For the truncated preprojective representations of $K(n)$ this is Theorem~\ref{thm:truncpp}. For the truncated preprojective representations lifted to the iterated universal abelian covering quivers $\widehat{K(n)}^{(k)}$ for $k\geq 1$, it follows inductively - when applying reflection recursions similar to those stated in (\ref{eq:recursive covers}) -  that these lifts are exceptional representations of $\widehat{K(n)}^{(k)}$, which means that their quiver Grassmannians are again smooth. Note that this is indeed clear for $m= 2$.

  The first part now follows by combining the results of Section~\ref{torusaction} and Theorems~\ref{thm:bb} and~\ref{cellscover}, taking into account that every preprojective representation of $\widetilde{ K(n)}$ is a lift of a preprojective representation (c.f.\ Lemma~\ref{le:lifts of transjectives}).
  The dual version for preinjective representations follows immediately since $\Gr_{\bfe}(P_m)\cong \Gr_{\udim P_m-\bfe}(I_m)$ and $\Gr_{\tbfe}(\tilde P_m)\cong\Gr_{\udim \tilde P_m-\tbfe}(\tilde I_m)$.

  The second part follows in the same way taking into account the initial remark.
\end{proof}

@@

\begin{remark}
  Below we will usually apply Lemma~\ref{lem:2equivalence} after performing each of the equivalences from Definition~\ref{def:2equivalence}.
  That is, given a 2-arrow $V:\Gamma(1)\implies\Gamma(2)$ for which $\Gamma(1)$ has a unique source $p$ and $\Gamma(2)$ has a unique sink $q$, we get an equivalent 2-quiver by replacing this 2-arrow with a usual arrow $\alpha_V:p\to q$.
\end{remark}
